The Epistemic Modal \emph{probable}

### 2019-03-24 paper.tex
formal part in beta mode, now write introduction

\documentclass{article}
\title{The Epistemic Modal \emph{probable}}
\author{Conrad Friedrich \\ MCMP, LMU Munich }
\date{\today}

\usepackage[utf8]{inputenc}

@@

\newcommand{\succeqK}{\succeq^K_{M,w}}
\begin{document}

@@

\todo{
  \begin{itemize}[nosep]
    \item Write an Introduction
     \item be perfectly clear as to what happens. SIGNPOST 
  \end{itemize}
}
\section{Intensional Semantics}
\todo{ \begin{itemize}[nosep]
  \item Tie the paper into framework used in class
\end{itemize} }

Note that we don't need first order logic. 
Abstracted away from syntax, just look at the semantics.
NEAT summary in Ivano's script, Class 9, page 3/4
\section{The Epistemic Modal \emph{probable}}
\todo{\begin{itemize}[nosep]
    \item  Give some natural language motivation
    \item Also show the need for expanding the standard intensional semantics!
    \item How is probable to be evaluated? Epistemic interpretation? MacFarlane's Solipcism.
    \item Mention how Lassiter has a different, graded adjective based approach.
    \item Discuss how absolute probability is best reduced to relative likelihood with the negation
      \item note that there are a lot more interesting problems associated with probable but to have an readable paper need to focus on particulars.
  \end{itemize}
}

To base the following discussions on a common language, we define a
propositional language $\L$ like \textcite{harrison-trainor17_prefer} and
\textcite{holliday13_measur} do:
\[
\varphi ::= p ~|~ \neg \varphi ~|~ (\varphi \land \psi) ~|~  (\varphi \succeq \psi) ~|~ \triangle\varphi ~|~ \Diamond \varphi
\]
Where $p \in \text{Atom}$, i.e. an atomic formula. The other boolean connectives $(\varphi \rightarrow \psi), (\varphi \vee
\psi)$ as well as $\top, \bot$ are defined in the standard way from negation and conjunction.
Define further $\Box \varphi = \neg\Diamond\neg \varphi$.
The intended readings are as follows:
\begin{align*}
  \varphi \succeq \psi &: \varphi \text{ is at least as likely as } \psi. \\
  \triangle \varphi &:\text{probably } \varphi. \\ 
                      \Diamond \varphi &: \text{It might be that } \varphi.
\end{align*}

\subsection{Criteria for Adequacy}
In his \parencite*{yalcin10_probab_operat}, Yalcin critically evaluates Kratzer's relative likelihood account and also Hamblin's account, which is based on the notion of possiblity (resp. plausiblity) functions \parencite{hamblin59_modal_probab}.
Yalcin does this by following the good philosophical practice of establishing a set of desiderata the accounts can be evaluated against. In the case for judging semantics for a natural language phenomenon, these naturally come in the form of inference patterns that strike one as intuitively valid or invalid.
The practice is fruitful because it gives criteria to compare different accounts in a transparent and comprehensible fashion, and shifts the focus of the problematic notion of \emph{intuitive plausibility} from judging the accounts themselves to judging the plausibility of the inference patterns.
Whether some formally specified semantics validates a given inference pattern is then simply not a matter of debate and can be shown mathematically given the assumptions one is forced to make explicit.   
The rules are not meant to characterize a logic, of course.
\todo{
\begin{itemize}[nosep]
  \item Motivate and describe the conditions
\end{itemize}
}
\begin{center}
    \begin{prooftree}
        \hypo{ \triangle \varphi } \infer1[(V1)]{ \neg \triangle \neg \varphi}
    \end{prooftree}
\end{center}

@@

\begin{center}
\begin{prooftree}
    \hypo{ \Box \varphi } \infer1[(V6)]{ \triangle \varphi}
    \end{prooftree}
\end{center}

\begin{center}
\begin{prooftree}
    \hypo{ \triangle \varphi } \infer1[(V7)]{ \Diamond \varphi}
    \end{prooftree}
\end{center}

@@

\section{Kratzer's Preferential Semantics}
\todo{
  \begin{itemize}
    \item Clean up this section a bit
  \end{itemize}
}
For \textcite{kratzer91_modal} epistemic modals like \emph{probably} are just a special case of a more general class of modal auxiliaries.
These can be epistemic (e.g. `Jockl must have been the murderer [in view of the available evidence]'), deontic (e.g. `Jockl must go to jail [in view of what the law provides]'), or circumstancial (e.g. `Jockl can lift the rock [given the weight of the rock and the conditions of Jockl's muscles, etc.]').
Further, there are teleological (in view of certain aims), bouletic (in view of certain wishes), and doxastic (in view of certain beliefs) modalities.
This list is by no means complete.

Kratzer postulates that modality is always relative to a certain kind of interpreting the modalities involved, following Peirce (cite).
She set out to formulate these different modalities in a common framework such that we'd only need to specify some parameters of the framework to determine the relevant modalities and give truth conditions for sentences involving modalities, while focussing on epistemic modals.

@@

More relevantly, the account for modality according to Kratzer needs to include graded modalities, of the following sort:
\begin{enumerate}[nosep,label=\alph*)]
  \item Michl must be the murderer.
  \item Michl is probably the murderer.
  \item There is a good possibility that Michl is the murderer.
  \item Michl might be the murderer.
\end{enumerate}
Here, the modalities are listed in decreasing strength, such that the intended reading is that a) implies b), b) implies c), and c) implies d).
Note that is a qualitative approach to graded modals, as opposed to quantitative approaches, employing a scale structure in the semantics, as described by \textcite{lassiter10_gradab}.

\todo{\begin{itemize}[nosep]
  \item The account presented has a couple of advantages.
\item Neatly fits with the other modals, just change f and g
\item Gives an account of probable without resorting to probability functions (give argument by Hamblin - maybe more in philosophy texts? check titelbaum)
\end{itemize}
}
How do the semantics look like?
We assume a set $W$ of possible worlds, a set of propositions $P$ such that elements of $P$ are subsets of $W$, and assume that a proposition $p$ is true at $w$ iff $w \in p$.
The crucial element in Kratzer's semantics that distinguishes her account from most simple semantics via Kripke frames with an accessibility relation is the \emph{conversational background}.
It has two components.

The \emph{modal base} determines for every world the set of worlds which are epistemically accessible from it.
More formally, the modal base is a function $f: W \rightarrow \wp(W)$ assigning each world $w$ a set of worlds.
The set of worlds $f(w)$ assigned to $w$ represents, in the epistemic case, what we know at $w$.

The \emph{ordering source} is a function $g: W \rightarrow \wp(W)$ s.t. each $w$ is assigned a set of propositions $g(w)$ which interpreted as an \emph{ideal} (cf. \cite{lewis81_order_seman_premis_seman_count}). The ordering source induces an order $\leq_{g(w)}$ on $W$ for all $w,w' \in W$ as follows:
\begin{equation}
    \label{eq:porder}
w \leq_{g(w)} w' :\iff \{p: p\in g(w) ~\&~ w' \in p\} \subseteq \{p: p\in  g(w) ~\&~ w \in p\}
\end{equation}
\todo{THIS NEED UPDATING SINCE g(w) CHANGED!}
\begin{equation}
    \label{eq:porder}
w \leq_{g(w)} w' :\iff \{A: g(w) \subseteq A ~\&~ w' \in A\} \subseteq \{A: g(w) \subseteq A ~\&~ w \in A\}
\end{equation}
The intended reading is that $w$ is at least as close to the ideal represented by $g(w)$ as $w'$ iff all the ideal propositions in $g(w)$ which are true in $w'$ are also true in $w$.
From this, Kratzer defines a the semantics for a list of different expressions, like `good possibility', `possibility' and so on.
For our purposes, it suffices to look at those definitions involved in defining the meaning of \emph{probably $\alpha$}.

First, define `as least as good a possibility as'.
The induced order on worlds $\leq_{g(w)}$ is `lifted' to an order on the level of propositions.
For a modal base $f$ and a ordering source $g$, for propositions $A,B$ and world $w$, we can define
\begin{equation}
    \label{eq:lifted}
    A \succeq_w B :\iff \forall b \in B_w \exists a \in A_w: a \leq_{g(w)} b,
\end{equation}
where $A_w = f(w) \cap A$, and $B_w$ analogously (cf. \cite[][p.~519]{holliday13_measur}).
In words, a proposition $A$ is as least as good a possibility as a proposition $B$ seen from a world $w$ iff for any world in $B$ epistemically accessible from $w$ there is a world in $A$ accessible from $w$ that is at least as close to the ideal picked out by $g(w)$.

The strict part is defined as one would expect:

\noindent $A \succ_w B :\iff {A \succeq_w B ~\&~ B \not\succeq_w A}$.

Finally, we can assign a meaning to \emph{probably $\alpha$} given a set of worlds $W$, a modal base $f$ and a ordering source $g$ \parencite[][p.~645]{kratzer91_modal}:
\begin{equation}
\llbracket \text{probably } \alpha \rrbracket^{f,g,w} = 1 :\iff \lb \alpha \rb \succ_w \lb \neg\alpha \rb
\end{equation}
Assuming here that we are looking at a flat fragment and $\alpha$ doesn't contain modals itself.
Intuitively,... \todo{[briefly explain]}

@@

To put it slightly more formally, we define: 
\begin{definition}
    A \emph{Kratzer preferential model} is a tuple $M = \langle W,f,g,V \rangle$ s.t.
    \begin{itemize}[nosep]
\renewcommand\labelitemi{--}
      \item $W$ is a non-empty set of possible worlds,
      \item $f,g$ are functions $f,g: W \rightarrow \wp(W)$,
      \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$
    \end{itemize}
\end{definition}

Further, we want the induced order defined above also hold for $M$.
This is done in the following uninteresting definition:
Let $f$ be a modal base, $g$ be a ordering source, $W$ a set of worlds.
For each $w \in W$ let $\succeq_w$ be the order on subsets of $W$ defined in
Equation \ref{eq:lifted}.
Let $M = \langle W,f,g,V \rangle$ be a Kratzer preferential model.
For $p,q \subseteq W$ we define an induced order on $\wp(W)$:   
\begin{equation}
    \label{eq:induced}
       p \succeq^{M,w} q \text{ iff } p \succeq_w q.
\end{equation}

@@

    \begin{enumerate}[nosep]
      \item If $p \in \text{Atom}$, $M,w \models p$ iff $w \in V(p)$,
      \item $M,w \models \neg \varphi$ iff $M,w \not\models \varphi$
      \item $M,w \models (\varphi \land \psi)$ iff $M,w \models \varphi$ and $M,w \models \psi$,
      \item $\lb \varphi \rb^M = \{w' \in W : M,w' \models \varphi\}$,
      \item $M,w \models \varphi \succeq \psi$ iff $\lb \varphi \rb^M \succeq^{M,w} \lb \psi \rb^M$,
      \item $M,w \models \triangle \varphi$ iff $M,w \models \varphi \succeq \neg \varphi ~\&~ \neg (\neg \varphi \succeq \varphi)$,
      \item $M,w \models \Diamond \varphi$ iff ...
    \end{enumerate}
\end{definition}

\todo{FORMALIZE MIGHT PHI}
\subsection{Discussion}

How does Kratzer's account fare with respect to the adequacy conditions we defined?
The semantics validate conditions (V1-V7), which can be easily seen by applying the definitions.
For example, let us establish that Kratzer preferential semantics satisfy (V2), i.e.: 
\begin{center}
\begin{prooftree}
        \hypo{ \triangle (\varphi \land \psi)} \infer1[(V2)]{ \triangle \varphi \land \triangle \psi}
    \end{prooftree}
\end{center}
Suppose $M,w \models \triangle (\varphi \land \psi)$.
Then $\lb (\varphi \land \psi) \rb^M \succ^{M,w} \lb \neg(\varphi \land \psi)\rb^M$.
Let $y \in \lb \neg \varphi \rb^M$.
Then $y \in \lb \neg \varphi \lor \neg\psi \rb^M = \lb \neg(\varphi \land \psi)\rb^M$.
By assumption, there is an $x \in \lb (\varphi \land \psi) \rb^M$ such that $x \leq_{g(w)} y$.
But then $x \in \lb \varphi \rb^M$, too.
Hence $\lb \varphi \rb^M \succ^{M,w} \lb \neg \varphi \rb^M$ and thus $M,w \models \triangle \varphi$.
The case for $M,w \models \triangle \psi$ works analogously.

Kratzer's account does \emph{not} validate (V11-V12).

\todo{THE FOLLOWING IS NOT CORRECT AND NEEDS TO BE REDONE}
\textcite[pp.~935, note~8]{yalcin10_probab_operat} gives a countermodel. In addition, the account \emph{does} validate pattern (I1), as can easily be seen. This is already an unwelcome result. Additionally, (I1) together with (V5) entails (I2) \parencite[p.~922]{yalcin10_probab_operat}. That is, on this account, anything that is probable is a least as likely as any tautology. This is, of course, a ruinous result.

@@

\noindent Let $\F$ be a boolean algebra. A possibility function is a function $p: \F \rightarrow [0,1]$ such that
\begin{enumerate}[nosep]
  \item $p(\emptyset) = 0$,
  \item $p(W) = 1$, and
  \item $p(U \cup V) = \max(p(U),p(V))$.
\end{enumerate}
\begin{definition}
    A \emph{Hamblin possibility model} is a tuple $M = \langle W,R,V,\mathcal{F},\Po\rangle$ such that
    \begin{itemize}[nosep]
        \renewcommand\labelitemi{--}
      \item $W$ is a set of possible worlds,
      \item $R$ is a binary relation on $W$,
        \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$,
        \item $\mathcal{F}$ is a function assigning each world $w \in W$ a boolean algebra $\F_w$ on $R(w)$, and 
          \item $\Po$ is a function assigning each world $w \in W$ a possibility function \\ $p_w: \F_w \rightarrow [0,1]$.
    \end{itemize}
\end{definition}

\begin{definition}[Hamblin possiblity semantics]
    \label{def:hambsem}
    Let $M = \langle M,R,V,\mathcal{F},\Po \rangle$ be a Hamblin possiblity model, $w \in W$, $\varphi, \psi \in \L$. We define clauses 1-4 and 6 just like in Kratzer preferential semantics (Definition \ref{def:kratsem}). We define further:
    \begin{enumerate}[nosep]
  \setcounter{enumi}{4}
      \item $M,w \models \varphi \succeq \psi$ iff $p_w(\lb \varphi \rb^M) \geq p_w(\lb \psi \rb^M) $,
  \setcounter{enumi}{6}
  \item $M,w \models \Diamond \varphi$ iff there is a $w' \in R(w)$ such that $ M,w' \models \varphi$.
    \end{enumerate}
\end{definition}
\noindent Here, we added a the standard semantics for $\Diamond \varphi$, as Yalcin proposes.
That the model contains a function assigning each world another function might seem formal overkill at first, but is just there such that the possiblity judgments are properly world-relative. When focusing on other aspects and e.g. developing a logic for this semantics, we might simplify our model in this regard, as \textcite{harrison-trainor17_prefer} do with preferential and probability models.

@@

Hamblin's account satisfies (V1-V7), as can easily be seen. For example, let us briefly validate (V2) like for Kratzer's account.
Suppose $M,w \models \triangle(\varphi \land \psi)$, hence $p_w(\lb(\varphi \land \psi )\rb^M) > p_w(\lb \neg (\varphi \land \psi) \rb^M)$.
Because $p_w(\lb (\varphi \land \psi)\rb^M \cup {\lb\neg(\varphi \land \psi)\rb^M)} = 1$, by (2) and (3) of the definition of possibility functions, we have $p_w(\lb (\varphi \land \psi) \rb^M) = 1$.
If $W$ is finite, there is $w' \in \lb (\varphi \land \psi) \rb^M$ such that $p_w(\{w'\}) = 1$. Since $w'$ is in $\lb \varphi \rb^M$ and $\lb \psi \rb^M$, but there is no $w'' \in \lb \varphi \rb^M\setminus\lb \psi \rb^M$ or $\lb \psi \rb^M\setminus\lb \varphi \rb^M$ with $p_w(\{w''\}) = 1$ we have $p_w(\lb \varphi \rb^M) > p_w(\lb \neg \varphi \rb^M)$ and $p_w(\lb \psi \rb^M) > p_w(\lb \neg \psi \rb^M)$ and hence $M,w \models \triangle \varphi \land \triangle \psi$.

@@

\begin{definition}
    A \emph{Yalcin probability model} is a tuple $M = \langle W,R,V,\mathcal{F},\Pr\rangle$ such that
    \begin{itemize}[nosep]
        \renewcommand\labelitemi{--}
      \item $W$ is a set of possible worlds,
      \item $R$ is a binary relation on $W$,
        \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$,
        \item $\mathcal{F}$ is a function assigning each world $w \in W$ a boolean algebra $\F_w$ on $R(w)$, and 
          \item $\Pr$ is a function assigning each world $w \in W$ a probability function \\ $\mu_w: \F_w \rightarrow [0,1]$.
    \end{itemize}
\end{definition}

\begin{definition}[Yalcin probability semantics]
    \label{def:yalcsem}
    Let $M = \langle M,R,V,\mathcal{F},\Pr \rangle$ be a Yalcin probability model, $w \in W$, $\varphi, \psi \in \L$. We define clauses 1-4 just like in Kratzer preferential models (Definition \ref{def:kratsem}). We define further:
    \begin{enumerate}[nosep]
  \setcounter{enumi}{4}
      \item $M,w \models \varphi \succeq \psi$ iff $\mu_w(\lb \varphi \rb^M) \geq \mu_w(\lb \psi \rb^M) $,
      \item $M,w \models \triangle \varphi$ iff $M,w \models \varphi \succ^{M,w} \neg \varphi$ iff $ \mu_w(\lb \varphi \rb^M) > .5$, and
  \item $M,w \models \Diamond \varphi$ iff there is a $w' \in R(w)$ such that $ M,w' \models \varphi$.
    \end{enumerate}
\end{definition}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}

@@

\newcommand{\succeqK}{\succeq^K_{M,w}}
\newcommand\wordcount{\input{paper.sum}}

\title{The Epistemic Modal \emph{probable}}
\author{Conrad Friedrich \\ MCMP, LMU Munich \\ \wordcount words}
\date{\today}

\begin{document}

@@

\todo{
  \begin{itemize}[nosep]
    \item Write an Introduction
     \item be perfectly clear as to what happens. SIGNPOST 
  \end{itemize}
}

\subsection{Intensional Semantics}
In this paper, we will not be concerned with type theoretical considerations. It seems also that the relevant literature is not very keen on discussion that. One reason might be that it is rather simple. We just view the \emph{probable}-operator as an operator on sentences, such that we just add something along the lines of `if $\varphi:t$ then $\triangle \varphi:t$' to our type theory. The interesting developments for the purposes of this paper take place when we enrich the model for our calculus. As a further simplification, we will only be concerned with a simple propositional language to not add further complexity by considering first order models, as will be described in the next section.

\section{The Epistemic Modal \emph{probable}}
\todo{\begin{itemize}[nosep]
    \item  Give some natural language motivation
    \item Also show the need for expanding the standard intensional semantics!
    \item How is probable to be evaluated? Epistemic interpretation? MacFarlane's Solipcism.
    \item Mention how Lassiter has a different, graded adjective based approach.
    \item Discuss how absolute probability is best reduced to relative likelihood with the negation
      \item note that there are a lot more interesting problems associated with probable but to have an readable paper need to focus on particulars.
  \end{itemize}
}

To base the following discussions on a common language, we define a
propositional language $\L$ like \textcite{harrison-trainor17_prefer} and
\textcite{holliday13_measur} do:
\[
\varphi ::= p ~|~ \neg \varphi ~|~ (\varphi \land \psi) ~|~  (\varphi \succeq \psi) ~|~ \triangle\varphi 
\]
Where $p$ is an atomic formula, i.e. $p \in \text{Atom}$. The other boolean connectives $(\varphi \rightarrow \psi), (\varphi \vee
\psi)$ as well as $\top, \bot$ are defined in the standard way from negation and conjunction.
The intended readings are as follows:
\begin{align*}
  \varphi \succeq \psi &: \varphi \text{ is at least as likely as } \psi. \\
  \triangle \varphi &:\text{probably } \varphi. \\ 
\end{align*}

\subsection{Criteria for Adequacy}
In his \parencite*{yalcin10_probab_operat}, Yalcin critically evaluates Kratzer's relative likelihood account and also Hamblin's account, which is based on the notion of possiblity (resp. plausiblity) functions \parencite{hamblin59_modal_probab}.
Yalcin does this by following the good philosophical practice of establishing a set of desiderata the accounts can be evaluated against. In the case for judging semantics for a natural language phenomenon, these naturally come in the form of inference patterns that strike one as intuitively valid or invalid.
The practice is fruitful because it gives criteria to compare different accounts in a transparent and comprehensible fashion, and shifts the focus of the problematic notion of \emph{intuitive plausibility} from judging the accounts themselves to judging the plausibility of the inference patterns.
Whether some formally specified semantics validates a given inference pattern is then simply not a matter of debate and can be shown mathematically given the assumptions one is forced to make explicit.   
The rules are not meant to characterize a logic, of course.

\todo{
\begin{itemize}[nosep]
  \item Motivate and describe the conditions
  \item Describe what inference means (truth preservation)
\end{itemize}
}
\begin{center}
    \begin{prooftree}
        \hypo{ \triangle \varphi } \infer1[(V1)]{ \neg \triangle \neg \varphi}
    \end{prooftree}
\end{center}

@@

@@

In addition, we weren't able to include formal considerations on how \emph{probable}-talk relates to other epistemic modals like \emph{might, must}, etc., due to time and place constraints, although this aspect is particularly lending itself to analysis and would be the first to consider in a longer paper.

\section{Kratzer's Preferential Semantics}
For \textcite{kratzer91_modal} epistemic modals like \emph{probably} are just a special case of a more general class of modal auxiliaries.
These can be epistemic (e.g. `Jockl must have been the murderer [in view of the available evidence]'), deontic (e.g. `Jockl must go to jail [in view of what the law provides]'), or circumstancial (e.g. `Jockl can lift the rock [given the weight of the rock and the conditions of Jockl's muscles, etc.]').
Further, there are teleological (in view of certain aims), bouletic (in view of certain wishes), and doxastic (in view of certain beliefs) modalities.
This list is by no means complete.

Kratzer postulates that modality is always relative to a certain kind of interpreting the modalities involved.
She sets out to formulate these different modalities in a common framework such that we'd only need to specify some parameters of the framework to determine the relevant modalities and give truth conditions for sentences involving modalities, while focusing on epistemic modals.

@@

More relevantly, the account for modality according to Kratzer needs to include graded modalities, of the following sort:
\begin{enumerate}[nosep,label=\alph*)]
  \item Michl must be the murderer.
  \item Michl is probably the murderer.
  \item There is a good possibility that Michl is the murderer.
  \item Michl might be the murderer.
\end{enumerate}
Here, the modalities are listed in decreasing strength, such that the intended reading is that a) implies b), b) implies c), and c) implies d).
Note that is a qualitative approach to graded modals, as opposed to quantitative approaches. Some of these employ a scale structure in the semantics, as described by \textcite{lassiter10_gradab}.

Kratzer intends her account to fit into a common framework for other gradable modals.
The idea is to just change the parameters $f,g$--to be introduced shortly--to represent the constraints that different modalities impose upon the semantics.
If successful, this idea would come out in favor of Kratzer's account, since it unifies multiple semantics into a single formal framework.
In addition, we'd have semantics for epistemic modals and \emph{probable}-talk that do not need to make recourse to probability theory.
Whether these two features actually do speak in favor of Kratzer's account shall not be evaluated in this paper and are just now mentioned.
We stick to the criteria defined in the beginning.

How do the semantics look like?
We assume a set $W$ of possible worlds, propositions are subsets of $W$, and assume that a proposition $p$ is true at $w$ iff $w \in p$.
The crucial element in Kratzer's semantics that distinguishes her account from most simple semantics via Kripke frames with only a single accessibility relation is the \emph{conversational background}.
It has two components.

The \emph{modal base} determines for every world the set of worlds which are epistemically accessible from it.
More formally, the modal base is a function $f: W \rightarrow \wp(W)$ assigning each world $w$ a set of worlds.\footnote{It is not perfectly clear from \textcite[][p.~644]{kratzer91_modal}, but it seems that the modal base in fact assigns a set of propositions to each world, so a set of sets of worlds. This is as far as we can tell equivalent for our purposes to the definition proposed here, if we make slight amendments.}
The set of worlds $f(w)$ assigned to $w$ is intended to represent what we know at $w$.

The \emph{ordering source} is a function $g: W \rightarrow \wp(W)$ s.t. each $w$ is assigned a set of propositions $g(w)$ which interpreted as an \emph{ideal} (cf. \cite{lewis81_order_seman_premis_seman_count}). The ordering source induces an ordering $\leq_{g(w)}$ on $W$ for all $w,w' \in W$ as follows:
% \begin{equation}
%     \label{eq:porder}
% w \leq_{g(w)} w' :\iff \{p: p\in g(w) ~\&~ w' \in p\} \subseteq \{p: p\in  g(w) ~\&~ w \in p\}
% \end{equation}
% \todo{THIS NEED UPDATING SINCE g(w) CHANGED!}
\begin{equation}
    \label{eq:porder}
w \leq_{g(w)} w' :\iff \{A: g(w) \subseteq A ~\&~ w' \in A\} \subseteq \{A: g(w) \subseteq A ~\&~ w \in A\}
\end{equation}
The intended reading is that $w$ is at least as close to the ideal represented by $g(w)$ as $w'$ iff all the ideal propositions in $g(w)$ which are true in $w'$ are also true in $w$.
From this, Kratzer defines a the semantics for a list of different expressions, like `good possibility', `possibility' and so on.
For our purposes, it suffices to look at the following ordering on propositions:

The induced ordering on worlds $\leq_{g(w)}$ is `lifted' to an ordering on the level of propositions.
For a modal base $f$ and a ordering source $g$, for propositions $A,B$ and world $w$, we can define
\begin{equation}
    \label{eq:lifted}
    A \succeq_w B :\iff \forall b \in B_w \exists a \in A_w: a \leq_{g(w)} b,
\end{equation}
where $A_w = f(w) \cap A$, and $B_w$ analogously (cf. \cite[][p.~519]{holliday13_measur}).
In words, a proposition $A$ is as least as good a possibility as a proposition $B$ seen from a world $w$ iff for any world in $B$ epistemically accessible from $w$ there is a world in $A$ accessible from $w$ that is at least as close to the ideal picked out by $g(w)$.

In the next section, we describe how this ordering can be used to define a semantics for our propositional language.

@@

\begin{definition}
    A \emph{Kratzer preferential model} is a tuple $M = \langle W,f,g,V \rangle$ s.t.
    \begin{itemize}[nosep]
\renewcommand\labelitemi{--}
      \item $W$ is a non-empty set of possible worlds,
      \item $f,g$ are functions $f,g: W \rightarrow \wp(W)$,
      \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$
    \end{itemize}
\end{definition}

Further, we want the induced order defined above also hold for $M$.
This is done via the following rather obvious definition:
Let $M = \langle W,f,g,V \rangle$ be a Kratzer preferential model.
For each $w \in W$ let $\succeq_w$ be the order on $\wp(W)$ defined in Equation \ref{eq:lifted}.
For $A,B \subseteq W$ we then define:   
\begin{equation}
    \label{eq:induced}
       A \succeq^{M,w} B \text{ iff } A \succeq_w B.
\end{equation}

@@

    \begin{enumerate}[nosep]
      \item If $p \in \text{Atom}$, $M,w \models p$ iff $w \in V(p)$,
      \item $M,w \models \neg \varphi$ iff $M,w \not\models \varphi$
      \item $M,w \models (\varphi \land \psi)$ iff $M,w \models \varphi$ and $M,w \models \psi$,
      \item $\lb \varphi \rb^M = \{w' \in W : M,w' \models \varphi\}$,
      \item $M,w \models \varphi \succeq \psi$ iff $\lb \varphi \rb^M \succeq^{M,w} \lb \psi \rb^M$,
      \item $M,w \models \triangle \varphi$ iff $M,w \models \varphi \succeq \neg \varphi ~\&~ \neg (\neg \varphi \succeq \varphi)$,
    \end{enumerate}
\end{definition}

How does Kratzer's account fare with respect to the adequacy conditions we defined?
The semantics validate conditions (V1-V5), which can be easily seen by applying the definitions.
For example, let us establish that Kratzer preferential semantics satisfy (V2), i.e.: 
\begin{center}
\begin{prooftree}
        \hypo{ \triangle (\varphi \land \psi)} \infer1[(V2)]{ \triangle \varphi \land \triangle \psi}
    \end{prooftree}
\end{center}
Suppose $M,w \models \triangle (\varphi \land \psi)$.
Then $\lb (\varphi \land \psi) \rb^M \succ^{M,w} \lb \neg(\varphi \land \psi)\rb^M$.
Let $y \in \lb \neg \varphi \rb^M$.
Then $y \in \lb \neg \varphi \lor \neg\psi \rb^M = \lb \neg(\varphi \land \psi)\rb^M$.
By assumption, there is an $x \in \lb (\varphi \land \psi) \rb^M$ such that $x \leq_{g(w)} y$.
But then $x \in \lb \varphi \rb^M$, too.
Hence $\lb \varphi \rb^M \succ^{M,w} \lb \neg \varphi \rb^M$ and thus $M,w \models \triangle \varphi$.
The case for $M,w \models \triangle \psi$ works analogously.

Kratzer's account does \emph{not} validate (V11-V12). \textcite[pp.~935, note~8]{yalcin10_probab_operat} gives a countermodel.
In addition, the account \emph{does} validate pattern (I1), as can easily be seen. This is already an unwelcome result. Additionally, (I1) together with (V5) entails (I2) \parencite[p.~922]{yalcin10_probab_operat}. That is, on this account, anything that is as probable as its negation is at least as likely as any tautology. This is, one can argue, a ruinous result.

Let's have a look at different accounts and see whether they fare better.

@@

\noindent Let $\F$ be a boolean algebra. A possibility function is a function $p: \F \rightarrow [0,1]$ such that
\begin{enumerate}[nosep]
  \item $p(\emptyset) = 0$,
  \item $p(W) = 1$, and
  \item $p(U \cup V) = \max(p(U),p(V))$.
\end{enumerate}
\begin{definition}
    A \emph{Hamblin possibility model} is a tuple $M = \langle W,R,V,\mathcal{F},\Po\rangle$ such that
    \begin{itemize}[nosep]
        \renewcommand\labelitemi{--}
      \item $W$ is a set of possible worlds,
      \item $R$ is a binary relation on $W$,
        \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$,
        \item $\mathcal{F}$ is a function assigning each world $w \in W$ a boolean algebra $\F_w$ on $R[w]$, and 
          \item $\Po$ is a function assigning each world $w \in W$ a possibility function \\ $p_w: \F_w \rightarrow [0,1]$.
    \end{itemize}
\end{definition}

\begin{definition}[Hamblin possiblity semantics]
    \label{def:hambsem}
    Let $M = \langle M,R,V,\mathcal{F},\Po \rangle$ be a Hamblin possiblity model, $w \in W$, $\varphi, \psi \in \L$. We define clauses 1-4 and 6 just like in Kratzer preferential semantics (Definition \ref{def:kratsem}). We define further:
    \begin{enumerate}[nosep]
  \setcounter{enumi}{4}
      \item $M,w \models \varphi \succeq \psi$ iff $p_w(\lb \varphi \rb^M) \geq p_w(\lb \psi \rb^M) $,
  \setcounter{enumi}{6}
    \end{enumerate}
\end{definition}
That the model contains a function assigning each world another function might seem formal overkill at first, but is just there such that the possiblity judgments are properly world-relative. When focusing on other aspects and e.g. developing a logic for this semantics, we might simplify our model in this regard, as \textcite{harrison-trainor17_prefer} do with preferential and probability models.

@@

Hamblin's account satisfies (V1-V5), as can easily be seen. For example, let us briefly validate (V2) like for Kratzer's account.
Suppose $M,w \models \triangle(\varphi \land \psi)$, hence $p_w(\lb(\varphi \land \psi )\rb^M) > p_w(\lb \neg (\varphi \land \psi) \rb^M)$.
Because $p_w(\lb (\varphi \land \psi)\rb^M \cup {\lb\neg(\varphi \land \psi)\rb^M)} = 1$, by (2) and (3) of the definition of possibility functions, we have $p_w(\lb (\varphi \land \psi) \rb^M) = 1$.
If $W$ is finite, there is $w' \in \lb (\varphi \land \psi) \rb^M$ such that $p_w(\{w'\}) = 1$. Since $w'$ is in $\lb \varphi \rb^M$ and $\lb \psi \rb^M$, but there is no $w'' \in \lb \varphi \rb^M\setminus\lb \psi \rb^M$ or $\lb \psi \rb^M\setminus\lb \varphi \rb^M$ with $p_w(\{w''\}) = 1$ we have $p_w(\lb \varphi \rb^M) > p_w(\lb \neg \varphi \rb^M)$ and $p_w(\lb \psi \rb^M) > p_w(\lb \neg \psi \rb^M)$ and hence $M,w \models \triangle \varphi \land \triangle \psi$.

@@

\begin{definition}
    A \emph{Yalcin probability model} is a tuple $M = \langle W,R,V,\mathcal{F},\Pr\rangle$ such that
    \begin{itemize}[nosep]
        \renewcommand\labelitemi{--}
      \item $W$ is a set of possible worlds,
      \item $R$ is a binary relation on $W$,
        \item $V$ is a function $V: \text{Atom} \rightarrow \wp(W)$,
        \item $\mathcal{F}$ is a function assigning each world $w \in W$ a boolean algebra $\F_w$ on $R[w]$, and 
          \item $\Pr$ is a function assigning each world $w \in W$ a probability function \\ $\mu_w: \F_w \rightarrow [0,1]$.
    \end{itemize}
\end{definition}

\begin{definition}[Yalcin probability semantics]
    \label{def:yalcsem}
    Let $M = \langle M,R,V,\mathcal{F},\Pr \rangle$ be a Yalcin probability model, $w \in W$, $\varphi, \psi \in \L$. We define clauses 1-4 just like in Kratzer preferential models (Definition \ref{def:kratsem}). We define further:
    \begin{enumerate}[nosep]
  \setcounter{enumi}{4}
      \item $M,w \models \varphi \succeq \psi$ iff $\mu_w(\lb \varphi \rb^M) \geq \mu_w(\lb \psi \rb^M) $,
      \item $M,w \models \triangle \varphi$ iff $M,w \models \varphi \succ^{M,w} \neg \varphi$ iff $ \mu_w(\lb \varphi \rb^M) > .5$, and
    \end{enumerate}
\end{definition}
